A thermodynamic view of dusty protoplanetary disks

### 2017-08-07 appendix.tex
resolve notation conflict in appendix C (now use curly K for the constant in drag law)
removed higher order correction section for submission, misc edits
misc home edits

\begin{align}
%  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
%  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
 \ii\sigma\frac{\Delta\rho}{\rho} &= \ii\sigma\frac{\delta\rho}{\rho} - 
  \dd\bm{v}\cdot\nabla\ln{\rho} =  \nabla\cdot\dd\bm{v},\label{lin_mass_full}\\
 %  \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
 % \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  \ii\sigma\frac{\Delta P}{P} &= \ii\sigma\frac{\delta P}{P} -
  \dd\bm{v}\cdot\nabla\ln{P} = 
\Gamma \nabla\cdot\dd\bm{v}  - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \hat{\bm{r}} \cdot \delta\bm{F} -  \hat{\bm{r}} \cdot \nabla\dd\psi ,\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \hat{\bm{z}} \cdot \delta\bm{F}  -  \hat{\bm{z}} \cdot \nabla\dd\psi, \\ 
\nabla^2 \delta\psi & = 4\pi G \dd\rho. 
\end{align}  
where the linearizd pressure force $\delta \bm{F}$ is given in
Appendix \ref{lin_press}. Recall $\Delta = \delta +
\bm{\xi}\cdot\nabla$ is the Lagrangian derivative and $\bm{\xi}$ is the
Lagrangian displacement. For axisymmetric flow, we have
$\ii\sigma\bm{\xi}\cdot\nabla = -\dd\bm{v}\cdot\nabla$.  These equations do not assume the
radially-local approximation used in numerical computations. Note that
Eq. \ref{lin_mass_full}---\ref{lin_energy_full} imply 
\begin{align}
\left|\sigma\right|^2\frac{\Delta P}{P}\frac{\Delta \rho^*}{\rho} =
\Gamma\left|\nabla\cdot\dd\bm{v}\right|^2 -
\nabla\cdot\dd\bm{v}^*\left(\dd\bm{v}\cdot\nabla\ln{K} + \frac{\dd\mathcal{C}}{P}\right),
\end{align}
so the first (real) term on the right hand side does not contribute to
$\imag{\left(\Delta P\Delta\rho^*\right)}$. Setting $K$ to constant
and taking the imaginary part gives Eq. \ref{pdv}.

@@

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop  \propto  \frac{1}{\rho c_s}\label{epstein}
\end{align}
\citep{price15}. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right) = - K c_s^2 \nabla\cdot
  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
  -Kc_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{Kc_s} = \bm{F}\cdot\nabla\dd\tepsilon +
  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
  \tepsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k_x -
    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k_x \right)\frac{Q}{c_s^2},\\
  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

@@

\begin{align}
  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\label{streaming_mass}\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\label{streaming_vx}\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\label{streaming_vz}\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
%{\bf bg pressure gradient neglected in energy eq (is this ok? could
%  fix, but get more complicated dispersion}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation. We note that the incompressible
condition used by \cite{jacquet11} is obtained by setting $\zeta$ to zero. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. \label{lin_drag_si}
\end{align}
%{\bf only the second term contributes to work done? dont think so}

@@

Analytic expressions for cubic roots are unwieldy. To keep the problem
tractable, let us \emph{assume} at this stage that the quadratic term
can be neglected compare to the last term. That is, 
%This amounts 
%to assuming a  sufficiently \emph{large} $K_z$ such that :
\begin{align}
\left|\frac{\ii\fdust}{2K_x\fg} + \left(1 -
      \fdust\right)\right| \left|K_x\nu\right|^2 \ll
  K_z^2\left(2\fdust - 1\right). \label{min_kz}
\end{align}
(For fixed $\fdust$ but allowing $K_x\to\infty$, the imaginary term can be
neglected.) We show in \S\ref{si_dust_rich_approx} that this
simplification is valid for dust-rich disks.

>>>>

\begin{align}
%  \ii\sigma\frac{\delta\rho}{\rho} &= \nabla\cdot\dd\bm{v} +
%  \dd\bm{v}\cdot\nabla\ln{\rho},\label{lin_mass_full}\\
 \ii\sigma\frac{\Delta\rho}{\rho} &= \ii\sigma\frac{\delta\rho}{\rho} - 
  \dd\bm{v}\cdot\nabla\ln{\rho} =  \nabla\cdot\dd\bm{v},\label{lin_mass_full}\\
 %  \ii\sigma\frac{\delta P}{P} &= \Gamma \nabla\cdot\dd\bm{v} +
 % \dd\bm{v}\cdot\nabla\ln{P} - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  \ii\sigma\frac{\Delta P}{P} &= \ii\sigma\frac{\delta P}{P} -
  \dd\bm{v}\cdot\nabla\ln{P} = 
\Gamma \nabla\cdot\dd\bm{v}  - \dd\bm{v}\cdot\nabla\ln{K} - \frac{\dd\mathcal{C}}{P}.\label{lin_energy_full}\\
  -\ii\sigma\dd v_r  &= 2\Omega\dd v_\phi + 
  \hat{\bm{r}} \cdot \delta\bm{F} -  \hat{\bm{r}} \cdot \nabla\dd\psi ,\\
  \ii\sigma\dd v_\phi &= \frac{\kappa^2}{2\Omega}\dd v_r + \frac{\p
    v_\phi}{\p z}\dd v_z,\\
  -\ii\sigma\dd v_z &=  \hat{\bm{z}} \cdot \delta\bm{F}  -  \hat{\bm{z}} \cdot \nabla\dd\psi, \\ 
\nabla^2 \delta\psi & = 4\pi G \dd\rho. 
\end{align}  
where the linearizd pressure force $\delta \bm{F}$ is given in
Appendix \ref{lin_press}. Recall $\Delta = \delta +
\bm{\xi}\cdot\nabla$ is the Lagrangian perturbation and $\bm{\xi}$ is the
Lagrangian displacement. In addition, 
$\ii\sigma\bm{\xi}\cdot\nabla = -\dd\bm{v}\cdot\nabla$ for axisymmetric flow.  These equations do not assume the
radially-local approximation used in numerical computations. Note that
Eq. \ref{lin_mass_full}---\ref{lin_energy_full} imply 
\begin{align}
\left|\sigma\right|^2\frac{\Delta P}{P}\frac{\Delta \rho^*}{\rho} =
\Gamma\left|\nabla\cdot\dd\bm{v}\right|^2 -
\nabla\cdot\dd\bm{v}^*\left(\dd\bm{v}\cdot\nabla\ln{K} + \frac{\dd\mathcal{C}}{P}\right),
\end{align}
so the first (real) term on the right hand side does not contribute to
$\imag{\left(\Delta P\Delta\rho^*\right)}$. Setting $K$ to constant
and taking the imaginary part gives Eq. \ref{pdv}.

@@

\section{Linearized dust diffusion}\label{lin_dust}
We consider small grains in the Epstein regime, with fixed internal
density and size, so that
\begin{align}
  \tstop   =   \frac{\mathcal{K}}{\rho c_s}\label{epstein}
\end{align}
\citep{price15}, where $\mathcal{K}$ is a constant. Then the dust diffusion
function becomes
\begin{align}
  \mathcal{C} \equiv c_s^2\nabla\cdot\left(\tepsilon\tstop\nabla
  P\right) = - \mathcal{K} c_s^2 \nabla\cdot
  \left(\frac{\tepsilon}{c_s}\bm{F}\right) =
  -\mathcal{K}c_s \left(\bm{F}\cdot\nabla\tepsilon + \tepsilon \nabla\cdot\bm{F}
  - \frac{1}{2}\tepsilon\bm{F}\cdot\nabla\ln{c_s^2}\right).  
\end{align}
Linearizing,
\begin{align}
  - \frac{\dd\mathcal{C}}{\mathcal{K}c_s} = \bm{F}\cdot\nabla\dd\tepsilon +
  \dd\bm{F}\cdot\nabla\tepsilon + \dd\tepsilon\nabla\cdot\bm{F} +
  \tepsilon\nabla\cdot\dd\bm{F} -
  \frac{1}{2}\nabla\ln{c_s^2}\cdot\left(\bm{F}\dd\tepsilon + \tepsilon
  \dd\bm{F}\right). 
\end{align}
The linearized dust-fraction $\dd\tepsilon$ and its derivatives in the
radially-local approximation are given by
\begin{align}
  \dd\tepsilon      &= (1-\tepsilon)W - \frac{Q}{c_s^2}, \\
  \p_r\dd\tepsilon &= \left[(1-\tepsilon)\left(\ii k_x -
    \p_r\ln{\rho}\right) - \p_r\tepsilon\right]W + \left(\p_r\ln{c_s^2}
  + \p_r\ln{\rho} - \ii k_x \right)\frac{Q}{c_s^2},\\
  \dd\tepsilon^\prime &= (1-\tepsilon)W^\prime - \tepsilon^\prime W -
  \left(\frac{Q}{c_s^2}\right)^\prime. 
\end{align}

@@

\begin{align}
  \ii \sigma W &=\nabla\cdot \dd \bm{ v} = \ii k_x \dd v_x + \ii k_z \dd v_z,\label{streaming_mass}\\
    \sigma^2 \dd v_x &= \kappa^2 \dd v_x - \ii \sigma F_r W +
    k_x\sigma Q,\label{streaming_vx}\\
  -\ii\sigma\dd v_z &= -\ii k_zQ,\label{streaming_vz}\\
\ii \zeta \sigma  Q & = \frac{P}{\rho} \nabla \cdot \dd\bm{v}   -
  \frac{\dd \mathcal{C}}{\rho}. 
\end{align}
%{\bf bg pressure gradient neglected in energy eq (is this ok? could
%  fix, but get more complicated dispersion}
The artifical factor $\zeta = 1$  is inserted to keep track of the
left-hand-side of the energy equation.  Setting $\zeta$ to zero is
equivalent to assuming incompressible gas \citep{jacquet11}. 
The linearized dust diffusion
function (Appendix \ref{lin_dust}), 
under the above approximations, is 
\begin{align}
-\frac{\dd \mathcal{C}}{\rho}  = \tstop c_s^2 \left[\ii k_x F_r\left( 1-2\tepsilon\right)
  W + \tepsilon |\bm{k}|^2 Q\right]. \label{lin_drag_si}
\end{align}
%{\bf only the second term contributes to work done? dont think so}

@@

Analytic expressions for cubic roots are unwieldy. To keep the problem
tractable, let us \emph{assume} at this stage that the quadratic term
can be neglected compare to the last term. That is, 
%This amounts 
%to assuming a  sufficiently \emph{large} $K_z$ such that :
\begin{align}
\left|\frac{\ii\fdust}{2K_x\fg} + \left(1 -
      \fdust\right)\right| \left|K_x\nu\right|^2 \ll
  K_z^2\left(2\fdust - 1\right). \label{min_kz}
\end{align}
(The imaginary term can be
neglected for fixed $\fdust$ but allowing $K_x\to\infty$.) We show in \S\ref{si_dust_rich_approx} that this
simplification is valid for dust-rich disks.
